---
title: "PaleoClimR: Features for Visualising and Analysing Palaeoclimate Output"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{PaleoClimR Features}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, message=FALSE, warning=FALSE}
library(PaleoClimR)

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Overview

This vignette demonstrates a basic workflow for importing, plotting and analysing palaeoclimate model output using PaleoClimR.

To keep package builds stable and fast, this example:

- uses a temporary directory
- avoids repeated downloads
- cleans up automatically

# Create a Temporary Working Directory

```{r}
tmp_dir <- file.path(tempdir(), "paleoclimr_example")
dir.create(tmp_dir, showWarnings = FALSE)

# ensure cleanup even if errors occur
on.exit(unlink(tmp_dir, recursive = TRUE), add = TRUE)

tmp_dir
```

# Download Example cGENIE Model Output

Large downloads should not run automatically during package checks.  

```{r}
url <- "http://www.seao2.info/cgenie_output/muffin.CB.p_worbe2.BASES.ridgwelletal.SPIN.tar.gz"
tarfile <- file.path(tmp_dir, "muffin.CB.p_worbe2.BASES.ridgwelletal.SPIN.tar.gz")

if (!file.exists(tarfile)) {
  download.file(url, tarfile, mode = "wb", method = "libcurl")
}

untar(tarfile, exdir = tmp_dir)
```

# Define cGENIE Experiment Path

```{r}
exp_path <- file.path(
  tmp_dir,
  "muffin.CB.p_worbe2.BASES.ridgwelletal.SPIN"
)

exp_path
```

# Map Visualization

```{r}

var <- "ocn_O2"

cGENIE.map(
  var = var,
  experiment = exp_path,
  depth.level = 1,
  dims = 3,
  year = "default",
  unit.factor = 1e6,
  min.value = 0,
  max.value = 300,
  intervals = 30,
  continents.outlined = TRUE,
  scale.label = "NULL",
  model = "biogem",
  line.thickness = 1,
  palette_name = pals::parula(1000),
  projection = "ESRI:54012",
  darkmode = FALSE
)
```
